---
title: "Chapter 01 - Data Visualization with ggplot2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
```{r}
library(tidyverse)
```

```{r}
mpg <- ggplot2::mpg
```

## Creating a ggplot
```{r}
# Exemplo no livro - p. 5
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point() # +
 # geom_smooth(method = "lm")
```

## A Graphing Template - Exercises (p. 6)

```{r}
#1
ggplot(data = mpg)
```

```{r}
#2 
nrow(mpg)
ncol(mpg)
# 234 linhas e 11 colunas
```

```{r}
# 3 
?mpg
# drv
# f = front-wheel drive, r = rear wheel drive, 4 = 4wd
```

```{r}
# 4
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

```{r}
# 5
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

Não é util porque nenhuma variável é contínua.

## Aesthetic Mappings

>"An aesthetic is a visual property of the objects in your plot. Aesthetics include thinks like the size, the shape, or the color of your points.


```{r}
# Exemplo. pg 8

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

> "Mapping an unordered variable (class) to an ordered aesthetic (size) is not a good idea.

```{r}
# Exemplo pg. 9
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```


```{r}
# Exemplo pg. 9
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```


```{r}
# Exemplo pg. 9
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```



```{r}
# Exemplo pg. 9
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```


## Aesthetic Mappings - Exercises (p. 12)

```{r}
#1
ggplot(data = mpg) + 
  geom_point(
    mapping = aes(x = displ, y= hwy), color = "blue"
  )
```

```{r}
# 2
glimpse(mpg)
# categorical: manufacturer, model, trans, drv, fl, class
# continuous: displ, year, cyk, cty, hwy
```


```{r}
# 3
# ggplot(data = mpg) +
#  geom_point(aes(x = displ, y = year, shape = cty))

#Erro: A continuous variable can not be mapped to shape
```

```{r}
# 4
 ggplot(data = mpg) +
  geom_point(aes(x = class, y = class, shape = class))

```

```{r}
# 5
 ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, stroke = cty))

```



```{r}
# 6
 ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = displ < 5))

```

## Facets

- é útil para variáveis categóricas discretas
- `facet wrap()` - para usar o facet com 1 variável, primeiro argumento seria uma formula.
- **formula** é o nome de uma estrutura de dados no R, não é sinônimo de equação
- `facet_grid()` - para plotar a combinação de duas variáveis. 


```{r}
# Ex. pg. 14
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

```{r}
# Ex. pg. 14
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```

```{r}
# Ex. pg. 15
mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  #facet_wrap(~ cyl)
  facet_grid(cyl ~ .)
```

## Facets - Exercises (p. 15)
```{r}
# 1

mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ cty)

# O resultado não é muito bom.. cria muitos gráficos
```

```{r}
# 2

mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)

# os gráficos vazios no caso acima representam gráficos onde não tem dados

mpg %>% 
  ggplot() +
  geom_point(mapping = aes(x = drv, y = cyl))

# são duas variáveis categóricas plotadas!
```


```{r}
# exercício 3

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)


ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)


# o ponto determina a dimensão, se vai plotar apenas gráficos verticais ou horizontais

```




```{r}
# exercício 4
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~class, nrow = 2)

# Separar com faceting facilita a visualização, principalmente se tiver muitos dados!!  
```

```{r}
# exercicio 5
# nrow define a quantidade de linhas verticais que o gráfico será separado, e ncol são as colunas.

#facet grid nao tem ncol e nrow pois as 2 variáveis ocupam as linhas e colunas


# ?facet_wrap
```

```{r}
# exercicio 6

# se tiver só uma categoria, não é necessário facet.
```


### Geometric objects

> A geom is the geometrical object that a plot uses to represent data. 


```{r}
# p. 16

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy))


ggplot(data = mpg) +
  geom_smooth(aes(x = displ, y = hwy))



ggplot(data = mpg) +
    geom_point(aes(x = displ, y = hwy, color = drv))+
  geom_smooth(aes(x = displ, y = hwy, linetype = drv, color = drv))

```

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy))+
  geom_point(mapping = aes(color = class)) +
  geom_smooth()
```

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy))+
  geom_point(mapping = aes(color = class)) +
  geom_smooth(
    data = filter(mpg, class == "subcompact"),
    se = FALSE
  )
```


#### Exercícios
```{r}
# 1

# line chart
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_line()

# boxplot
ggplot(data = mpg, aes(x = displ, y = hwy, group = class)) +
geom_boxplot()

# histogram
ggplot(data = mpg, aes(x = displ)) +
  geom_histogram()

# area chart
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_area()
```



```{r}
# 2 
ggplot(data = mpg, aes(x = displ, y = hwy, color = drv))+
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r}
# 3
# show.legend = FALSE, o gráfico fica sem legendas
```

```{r}
# 4. `se` argument é para aparecer ou não o intervalo de confiança
```

```{r}
#5. os gráficos serão iguais

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

```{r}
# 6
base <- 
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) 

# 1
base + 
  geom_point()+
  geom_smooth(se = FALSE)

# 2 
base + 
  geom_point()+
  geom_smooth(aes(fill = drv), se = FALSE)

# 3
base + 
  geom_point(aes(color = drv))+
  geom_smooth(aes(color = drv), se = FALSE)

# 4 
base + 
  geom_point(aes(color = drv))+
  geom_smooth(se = FALSE)

# 5
base + 
  geom_point(aes(color = drv))+
  geom_smooth(aes(linetype = drv), se = FALSE)


# 6
base + 
  geom_point(aes(color = drv))


```


## Statistical transformations

```{r}
ggplot(data = diamonds) + 
  geom_bar(aes(x = cut))
```

```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

```{r}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
```

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

### Exercícios

1) What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

2) What does geom_col() do? How is it different to geom_bar()?

3) Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

4) What variables does stat_smooth() compute? What parameters control its behaviour?

5) In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?


```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = 1))
```
