---
title: "10 - Tibbles"
output: html_document
---

> It’s difficult to change base R without breaking existing code, so most innovation occurs in packages.

Para saber mais sobre o pacote:

```{r}
vignette("tibble")
```

```{r}
library(tidyverse)
```

>  Tibbles are one of the unifying features of the tidyverse. 

```{r}
as_tibble(iris)
```

> "It’s possible for a tibble to have column names that are not valid R variable names, aka non-syntactic names."

Ex:

```{r}
tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)

tb
```

- tribble: transposed tibble.
ex:

```{r}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)

```

> There are two main differences in the usage of a tibble vs. a classic data.frame: printing and subsetting.
Tibbles are designed so that you don’t accidentally overwhelm your console when you print large data frames

```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```



> Compared to a data.frame, tibbles are more strict: they never do partial matching, and they will generate a warning if the column you are trying to access does not exist.


```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df %>%  .$x

df %>%  .[["x"]]
```


> Some older functions don’t work with tibbles. If you encounter one of these functions, use as.data.frame() to turn a tibble back to a data.frame:

> With tibbles, [ always returns another tibble.

## 10.5 Exercícios

1.How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame). 

```{r}
class(mtcars)
```


```{r}
class(nycflights13::flights)
```



2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a") # transforma o conteúdo de xyz em factor

df


df$x # não dá erro, mesmo a coluna x não existindo
df[, "xyz"]  # retorna o conteúdo como um vetor
df[, c("abc", "xyz")] # retorna um df
```
```{r}
df_tibble <- tibble(abc = 1, xyz = "a")
df_tibble

df_tibble$x # deixa claro q a coluna não existe

df_tibble[, "xyz"] # retorna o conteúdo como uma tibble

df_tibble[, c("abc", "xyz")] # retorna uma tibble
```

3.If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r}
var <- "mpg"

mtcars_tibble <- as_tibble(mtcars)


mtcars_tibble[, var]
```

4. Practice referring to non-syntactic names in the following data frame by:




```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

4.1 Extracting the variable called 1.
```{r}
# 4.1

annoying$`1`
```

4.2 Plotting a scatterplot of 1 vs 2.
```{r}
# 4.2

annoying %>% 
  ggplot() +
  geom_point(aes(x = `1`, y = `2`))
```

4.3 Creating a new column called 3 which is 2 divided by 1.
```{r}
# 4.3

annoying$`3` <- annoying$`2` / annoying$`1`

annoying
```

4.4 Renaming the columns to one, two and three.
```{r}
# 4.4

annoying <- annoying %>%  rename(c(one = `1`, two = `2`, three = `3`)) 


annoying
```


5. What does tibble::enframe() do? When might you use it? 

```{r}

x <- sample(1:10)

tibble::enframe(x) # enframe recebe um vetor (1 dimensão), e cria um tibble, sendo criado uma coluna chamada name com a ordem de cada conteúdo do vetor, e uma coluna chamada value, contendo os conteúdos do vetor. 
```


6. What option controls how many additional column names are printed at the footer of a tibble?

```{r}
nycflights13::flights %>% 
  print(n_extra  = 5) # não parece ter diferença..
```

