---
title: "15 Factors"
output: html_document
---

## Introduction 

- Usado para variáveis categóricas

## Prerequisites

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(forcats)
```

## Creating factors

```{r}

x1 <- c("Dec", "Apr", "Jan", "Mar")

x2 <- c("Dec", "Apr", "Jam", "Mar")


month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
```

```{r}
y1 <- factor(x1, levels = month_levels)

y1

y2 <- factor(x2, levels = month_levels)

y2
```

- Qualquer valor que não esteja nos levels, será SILENCIOSAMENTE transformado em NA. Silenciosamente = o R não vai avisar! Cuidado com erros de digitação.

- Use a função `parse_factor()` para ver os erros/transformações em NA.
```{r}
y2 <- parse_factor(x2, levels = month_levels)

```

- Se não especificar os levels, os mesmos serão criados a partir dos dados, em ordem alfabética:
```{r}
factor(x1)
```

- Para criar factors com levels na ordem que aparecem, pode usar:

```{r}
f1 <- factor(x1, levels = unique(x1))
f1


f2 <- x1 %>% factor() %>% fct_inorder()
f2

```

- Para ver os níveis, usar `levels()`

```{r}
levels(f2)
```

## General Social Survey


```{r}
gss_cat <- forcats::gss_cat
```

```{r}
gss_cat %>% count(race)
```

```{r}
gss_cat %>%  ggplot(aes(race)) +
  geom_bar()
```
- Por padrão, o ggplot2 descarta levels que não tem valores!
```{r}
gss_cat %>%  ggplot(aes(race)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE)
```

- As operações mais comuns trabalhando com factors são: alterar a ordem dos levels, ou modificar o conteúdo dos levels.

### Exercícios

1) Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?
```{r}
gss_cat %>% ggplot(aes(rincome)) +
  geom_bar()
```
É difícil de entender pois tem muitos levels, e eles aparecem um em cima do outro!

2) What is the most common relig in this survey?
```{r}
gss_cat %>% count(relig) %>% arrange(desc(n)) %>% dplyr::slice(1) 
```

What’s the most common partyid?

```{r}
gss_cat %>% count(partyid) %>% arrange(desc(n)) %>% dplyr::slice(1) 
```

3) Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?


```{r}
gss_cat %>% filter(
  !denom %in% c(
    "No denomination",
    "Not applicable",
    "No denomination",
    "No answer",
    "Don't know"
  )
) %>%
  filter(relig != "No answer") %>%
  group_by(relig) %>%
  count()
```


```{r}
gss_cat %>%
  ggplot() +
  geom_bar(aes(x = relig, fill = denom)) +
  coord_flip()
```


## Modifying Factor Order


```{r}
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

ggplot(relig_summary, aes(tvhours, relig)) + geom_point()
```

```{r}
ggplot(relig_summary, aes(tvhours, fct_reorder(relig, tvhours))) +
  geom_point()
```

```{r}
relig_summary %>%
  mutate(relig = fct_reorder(relig, tvhours)) %>%
  ggplot(aes(tvhours, relig)) +
    geom_point()
```

```{r}
rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

ggplot(rincome_summary, aes(age, fct_reorder(rincome, age))) + geom_point()
```

```{r}
ggplot(rincome_summary, aes(age, fct_relevel(rincome, "Not applicable"))) +
  geom_point()
```
```{r}
by_age <- gss_cat %>%
  filter(!is.na(age)) %>%
  count(age, marital) %>%
  group_by(age) %>%
  mutate(prop = n / sum(n))

ggplot(by_age, aes(age, prop, colour = marital)) +
  geom_line(na.rm = TRUE)

ggplot(by_age, aes(age, prop, colour = fct_reorder2(marital, age, prop))) +
  geom_line() +
  labs(colour = "marital")
```

```{r}
gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(marital)) +
    geom_bar()

```

### Exercises

1) There are some suspiciously high numbers in tvhours. Is the mean a good summary?

Não, pois a média é mais influenciada por outliers. A mediana seria melhor neste caso. 


2) For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.

Apenas a variável rincome tem levels ordenados. 

```{r}
gss_cat %>% select_if(is.factor) %>% sapply(levels)
```


3) Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?

Porque a ordem estava de "baixo para cima", ou seja, do primeiro level até o último.

## Modifying Factor Levels



```{r}
gss_cat %>% count(partyid)
```


```{r}
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat"
  )) %>%
  count(partyid)

```

```{r}
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat",
    "Other"                 = "No answer",
    "Other"                 = "Don't know",
    "Other"                 = "Other party"
  )) %>%
  count(partyid)

```

```{r}
gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  count(partyid)
```

```{r}
gss_cat %>%
  mutate(relig = fct_lump(relig)) %>%
  count(relig)

```

```{r}
gss_cat %>%
  mutate(relig = fct_lump(relig, n = 10)) %>%
  count(relig, sort = TRUE) %>%
  print(n = Inf)

```

### Exercises


1) How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?


```{r}
gss_cat %>%
  mutate(partyid_small = fct_collapse(
    partyid,
    Democrat = c("Not str democrat", "Strong democrat"),
    Republican = c("Strong republican", "Not str republican"),
    Independent = c("Ind,near rep", "Independent", "Ind,near dem")
  )) %>%
  filter(partyid_small %in% c("Democrat", "Republican", "Independent")) %>% 
  count(year, partyid_small) %>% 
 ggplot()+
  geom_col(aes(x = year, y= n))+ 
  facet_wrap(~ partyid_small)
```

2) How could you collapse rincome into a small set of categories?

```{r}
gss_cat %>% count(rincome)
```


```{r}
gss_cat %>% 
  mutate(rincome_small = 
           forcats::fct_collapse(rincome,
"$1000 to 4999" = c("$1000 to 2999",
                    "$3000 to 3999",
                    "$4000 to 4999"),

"$5000 to 9999" = c("$5000 to 5999", "$6000 to 6999", "$7000 to 7999", "$8000 to 9999"
  
))) %>% 
  count(rincome_small)
```

